\documentclass[12pt]{article}
\usepackage{amsmath, amssymb, geometry}
\geometry{margin=1in}
\title{Recursive Information Saturation Collapse Theorem (RISC)}
\author{}
\date{}

\begin{document}
\maketitle

\section*{I. Theorem Statement}

The RISC Theorem formalizes a collapse condition in symbolic systems that experience either entropy overload or recursive saturation. The model identifies two distinct failure modes: entropy accumulation and recursion density.

Let:
\begin{itemize}
  \item $\Phi(t)$ – Coherence function at time $t$
  \item $H(t)$ – Entropy imbalance (incoming entropy minus dissipated entropy)
  \item $D_r(t)$ – Recursion density (recursive operations per time unit)
  \item $\Theta$ – Entropy collapse threshold
  \item $\theta_c$ – Recursion collapse threshold
\end{itemize}

Collapse occurs if:

\[
H(t) > \Theta \quad \text{or} \quad D_r(t) > \theta_c
\]

At this point:

\[
\frac{d^2 \Phi}{dt^2} > 0 \quad \text{and} \quad \Delta I(t) \rightarrow 0
\]

\section*{II. Collapse Mechanism}

\begin{enumerate}
  \item \textbf{Entropy Saturation:} Symbolic input exceeds system’s dissipation capacity.
  \item \textbf{Recursive Overload:} Internal referencing increases symbolic load beyond structural limits.
  \item \textbf{Feedback Amplification:} Attempts at stabilization raise recursion density and worsen collapse risk.
\end{enumerate}

\section*{III. Collapse Indicators}

\begin{itemize}
  \item Increasing $\frac{d^2 \Phi}{dt^2}$ (coherence loss acceleration)
  \item Declining mutual information between recursive outputs
  \item Entropy growth (Shannon or cross-entropy)
  \item Repetition, contradiction, or nonsensical output
\end{itemize}

\section*{IV. Simulation Framework}

\subsection*{Applicable Systems}

\begin{itemize}
  \item Autoregressive language models
  \item Symbolic logic systems with recursive input
  \item Cognitive agents under symbolic overload
\end{itemize}

\subsection*{Metrics}

\begin{itemize}
  \item BERTScore, BLEU, or coherence metrics
  \item Mutual information via MINE or CLUB
  \item Cosine similarity of embeddings
  \item Entropy trends over recursion depth
\end{itemize}

\section*{V. Visualization}

\begin{itemize}
  \item Time-series plots of $\Phi(t)$, $H(t)$, and $D_r(t)$
  \item Collapse prediction based on $\frac{d^2 \Phi}{dt^2}$ and $\Delta I(t)$
  \item UMAP/t-SNE embedding trajectories over recursive steps
\end{itemize}

\section*{VI. Cognitive Analogy}

\begin{itemize}
  \item Symbolic overload in paradox resolution tasks
  \item Working memory collapse from recursive reasoning
  \item Contradiction loops in high-bandwidth symbolic cognition
\end{itemize}

\section*{VII. Applications}

\begin{itemize}
  \item AI safety through entropy-aware model architecture
  \item Cognitive collapse modeling under symbolic recursion
  \item Symbolic system design with recursive fail-safes
  \item Early-warning metrics for system saturation
\end{itemize}

\section*{VIII. Future Work}

\begin{itemize}
  \item Model entropy input/output balance as a function of recursion
  \item Implement dampening layers for recursion in AI architectures
  \item Generalize RISC to multi-agent or graph-based systems
\end{itemize}

\end{document}